---
title: "iteration and list columns"
author: "Jake Vettoretti"
date: "2025-10-28"
output: html_document
---

```{r}
library(tidyverse)
library(rvest)
```

## Make a list
```{r}
l = #creating a list
  list(
    vec_numeric = 1:23,
    char_vec = c("Jeff"),
    mat = matrix(1:8, nrow = 2, ncol = 4),
    summary(rnorm(1000, mean = 4))
  )

l

#pulling out items from list
l[[1]]
l[["vec_numeric"]]
#try to avoid using $, is for pulling things out of lists 
```

## Make a different list
```{r}
list_normals = #making new list with different contents
  list(
    a = rnorm(30, mean = 3, sd = 1),
    b = rnorm(30, mean = 30, sd = 1),
    c = rnorm(30, mean = 3, sd = 10),
    d = rnorm(30, mean = -3, sd = 4)
  )

list_normals
```

(copy and paste function for last time)
```{r}
source("source/mean_and_sd.R")
```


Use a for loop to iterate these lists by this function

```{r}
output = vector("list", length = 4)

for (i in 1:4){
  
  output[[i]] = mean_and_sd(list_normals[[i]])
  
}

output #get list of four dataframes, can keep track of inputs and outputs
```

Use `map` to do the same thing

```{r}
output = map(list_normals, mean_and_sd) #can switch out mean_and_sd function to apply whatever function you want
output = map(list_normals, median) #using median this time
output #map statements let you use various functions for lists and don't have to set length of output object, empty spots needed etc.
```

`Map` variants, this one collapses everything into a df
```{r}
map_dfr(list_normals, mean_and_sd, .id = "sample")

map_dbl(list_normals, median)
```

## List columns

Try to put my list into a dataframe.

```{r}
listcol_df = 
  tibble(
    name = c("a", "b", "c", "d"), #both rows have link 4 so can put them next to each other in df
    sample = list_normals
  )
```

Did this work? Test by pulling list

```{r}
pull(listcol_df, sample)
pull(listcol_df, name)
```

Can I apply mean_and_sd
```{r}
mean_and_sd(pull(listcol_df,sample)[[1]])
mean_and_sd(pull(listcol_df,sample)[[2]])
mean_and_sd(pull(listcol_df,sample)[[3]])
mean_and_sd(pull(listcol_df,sample)[[4]])
```

Iterate using map
```{r}
map(pull(listcol_df, sample), mean_and_sd)
```

Adding a column

```{r}
listcol_df = 
  listcol_df |> 
  mutate(
    summary = map(sample, mean_and_sd)
  )

pull(listcol_df, summary)
```

```{r}
listcol_df |> 
  select(-sample) |> 
  unnest(summary)
```

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

nsduh_import = function(html, table_num) {
  
  data = 
    html |> 
    html_table() |> 
    nth(table_num) |>
    slice(-1) |> 
    select(-contains("P Value")) |>
    pivot_longer(
      -State,
      names_to = "age_year", 
      values_to = "percent") |>
    separate(age_year, into = c("age", "year"), sep = "\\(") |>
    mutate(
      year = str_replace(year, "\\)", ""),
      percent = str_replace(percent, "[a-c]$", ""),
      percent = as.numeric(percent)) |>
    filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
  
  data
  
}

nsduh_import(nsduh_html, table_num = 1)
nsduh_import(nsduh_html, table_num = 2)
nsduh_import(nsduh_html, table_num = 3)
```

Try this with a for loop

```{r}

output = vector("list", length = 3)

for ( i in 1:3) {
  
  output[[i]] = nsduh_import(html = nsduh_html, i) #function now has two specified arguments
  
}

output

```


Do this with `map`
```{r}
map(1:3, nsduh_import, html = nsduh_html) #mapping across 1, 2, and 3 using new nsduh function
```

Trying to tidy up by doing this all in a dataframe

```{r}

nsduh_df = 
  tibble(
    name = c("marj year", "marj month", "marj first"),
    number = 1:3 
  ) |> 
  mutate(
    table = map(number, nsduh_import, html = nsduh_html) #map across number column, apply this function, and do it here
  ) |> 
  unnest(table) #expands the entire thing, can take away
nsduh_df
```

## Look at weather data

```{r}
library(p8105.datasets)
data("weather_df")
```

Let's do a regression
```{r}

```

Let's iterate differently...
```{r}
weather_df |> 
  nest(data = date:tmin)
```

